### 2024-09-23 manuscript.tex
Improve precip. detection section

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
a cloud by the cloud detection method used in the ALCF. This is undesirable if
equivalent precipitation backscatter is not included in the simulated lidar
profiles. It was not possible to include precipitation simulation in the ALCF
due to the absence of required fields in the ICON model output and the
reanalysis data, the liquid and ice precipitation mass mixing ratios, and
because the required radiation calculations for precipitation are not
implemented in the ALCF. In this situation, in order to achieve a fair
comparison of observations and models, one should exclude profiles with
precipitation either manually or using an automated method. It is relatively
difficult to distinguish precipitation backscatter from cloud backscatter in
lidar observations, especially when only one wavelength channel and no
polarisation channel are available. In models, the same can be accomplished
relatively easily by excluding profiles with an amount of surface precipitation
flux exceeding a given threshold. In the observations, using precipitation flux
measurements can be very unreliable on ships due to ship movement and
turbulence caused by nearby ship structures. Our analysis of rain gauge data
from the RV \emph{Tangaroa} showed large discrepancies between the rain gauge
time series and human-performed synoptic observations, as well as large
inconsistencies in the rain gauge time series. Therefore, it is desirable to
implement a method of detecting precipitation from observed backscatter
profiles instead.

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence and type. We
used this dataset to train a convolutional artiﬁcial neural network (ANN) of
the U-Net type \citep{ronneberger2015} to recognise profiles with precipitation
from lidar backscatter. Samples of short time intervals (10 min) of
near-surface backscatter (0–250 m) were classified as clear, rain, snow, and
fog, using the synoptic observations as a training dataset. From these, a
binary classification of profiles as wet (rain or fog) and dry (clear or fog)
was derived.

The ANN achieved 65\% sensitivity and 87\% speciﬁcity when the true positive
rate (26\%) was made to match observations. We considered these success rates
satisfactory for the purpose of filtering precipitation profiles.

>>>>

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/ann.pdf}
\caption{
Artificial neural network (ANN) for prediction of precipitation in lidar
backscatter. \textbf{(a)} Diagram showing the TensorFlow structure of the ANN,
\textbf{(b)} randomly selected example samples of near-surface backscatter in
four categories (clear, fog, rain and snow), as determined by coincident
human-performed weather observations, \textbf{(c)} receiver operating
characteristic diagram of the ANN, \textbf{(d)} examples of 10-day time series
of human-observed (`HUM') and predicted precipitation based on an ANN trained
on all voyages (`ANN') and all voyages except for the shown voyage (`ANN2')
during three randomly selected voyages with the available data. Here, by
`randomly selected' we mean selected from the top of a permutation generated by
a pseudo-random number generator to prevent authors' bias in the selection.
}
\label{fig:ann}
\end{figure}

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
a cloud by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios), and because the required radiation calculations for precipitation are
currently not implemented in the ALCF. In order to achieve a fair comparison of
observations with models, one should exclude observed and simulated lidar
profiles with precipitation either manually or using an automated method. It is
relatively difficult to distinguish precipitation backscatter from cloud
backscatter in lidar observations, especially when only one wavelength channel
and no polarised channel are available. In models, the same can be accomplished
relatively easily by excluding profiles exceeding a certain amount of surface
precipitation flux. In the observations, using precipitation flux measurements
from rain gauges can be very unreliable on ships due to ship movement and
turbulence caused by nearby ship structures. Our analysis of rain gauge data
from the RV \emph{Tangaroa} showed large discrepancies between the rain gauge
time series and human-performed synoptic observations, as well as large
inconsistencies in the rain gauge time series. Human-performed observations of
precipitation presence or absence are expected to be reliable, but only cover a
limited set of time instants. Therefore, it was desirable to implement a method
of detecting precipitation from observed backscatter profiles alone.

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence or absence and
type. We used this dataset to train a convolutional artiﬁcial neural network
(ANN) of the U-Net type \citep{ronneberger2015} to recognise profiles with
precipitation from lidar backscatter (Fig. \ref{fig:ann}a), implemented in the
TensorFlow ANN framework \citep{tensorflow}. Samples of short time intervals
(10 min) of near-surface lidar backscatter (0–250 m) were classified as clear,
rain, snow, and fog, using the synoptic observations as a training dataset
(Fig.  \ref{fig:ann}b). From these, a binary, mutually exclusive classification
of profiles as precipitating (rain or snow) or dry (clear or fog) was derived.
For detecting model and reanalysis precipitation, we used a fixed threshold for
surface precipitation flux of 0.1 mm h$^{-1}$ (the ANN was not used).

The ANN achieved 65\% sensitivity and 87\% speciﬁcity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations.
